---
title: "RepData_PeerAssessment1"
author: "David Manteigas"
date: "Tuesday, April 14, 2015"
output: html_document
---
#Introduction

This assignment was performed using the [Activity Monitoring Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip). This dataset contains  data from a personal activity monitoring device.
The variables present in this dataset represent:
-**steps**: Number of steps taking in a 5-minute interval (missing values are coded as NA)
-**date**: The date on which the measurement was taken in YYYY-MM-DD format
-**interval**: Identifier for the 5-minute interval in which measurement was taken

All the calculations were performed using R x64 3.1.1 and/or confirmed by RStudio

##Loading and preprocessing the data

For this part of the assignment, the data was loaded into R and only the variable data was transformed into "Date" format.

```{r}
Activity_data <- read.csv("C:/Users/David Manteigas/Documents/activity.csv",stringsAsFactors=FALSE)
Activity_data$date <- as.Date(Activity_data$date)
str(Activity_data)
```
##What is mean total number of steps taken per day?

For this part of the assignment, the package dplyr will be used.
The following code transforms Activity_data into a tbl and then calculates the total number of steps by day.
```{r}
library(dplyr)
Activity_data_tbl <- tbl_df(Activity_data[,1:2])
Sum_steps_day <- Activity_data_tbl %>% group_by(date) %>% summarise_each(funs(sum(.,na.rm=TRUE)))
print(Sum_steps_day)
```

Here an histogram of the total steps by day is produced.
```{r}
with(Sum_steps_day,{
        hist(steps,col="blue",xlab="Total number of steps by day",ylab="Frequency")
	}
)
```

Mean and Median steps taken by day, using dplyr

```{r}
Mean_steps_day <- Activity_data_tbl %>% group_by(date) %>% summarise_each(funs(mean(.,na.rm=TRUE)))
Median_steps_day <- Activity_data_tbl %>% group_by(date) %>% summarise_each(funs(median(.,na.rm=TRUE)))
print(Mean_steps_day)
print(Median_steps_day)
```

##What is the average daily activity pattern?
Time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis). First, using dplyr and subsetting by day and interval and than using aggregate and filtering only by interval.
```{r}
Activity_data_tbl <- tbl_df(Activity_data)
Mean_steps_day_by_interval <- Activity_data_tbl %>% group_by(interval,date) %>% summarise_each(funs(mean(.,na.rm=TRUE)))
plot(steps~interval,data=Mean_steps_day_by_interval,type="l",ylab="Mean steps by day",xlab="Day interval")
stepsInterval <- aggregate(steps ~ interval, data = Activity_data, mean, na.rm = TRUE)
plot(steps ~ interval, data = stepsInterval, type = "l")
```

Calculation of the maximum number of steps taken by day and interval
```{r}
max_num_steps <- max(Mean_steps_day_by_interval$steps,na.rm=TRUE)
print(max_num_steps)
```

##Imputing missing values

Calculating the total number of NA's in the dataset
```{r}
total_na <- sum(is.na(Activity_data$steps))
```

Fill all the NA's in the dataset
```{r}
interval2steps <- function(interval) {
    stepsInterval[stepsInterval$interval == interval, ]$steps
} #made a function "interval2steps" to get the mean steps for particular 5-minute interval
activityFilled <- Activity_data  # Make a new dataset with the original data
count = 0  # Count the number of data filled in
for (i in 1:nrow(activityFilled)) {
    if (is.na(activityFilled[i, ]$steps)) {
        activityFilled[i, ]$steps <- interval2steps(activityFilled[i, ]$interval)
        count = count + 1
    }
}
cat("Total ", count, "NA values were filled.\n\r")
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
totalSteps2 <- aggregate(steps ~ date, data = activityFilled, sum)
hist(totalSteps2$steps)
mean(totalSteps2$steps)
median(totalSteps2$steps)
```

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
activityFilled$day = ifelse(as.POSIXlt(as.Date(activityFilled$date))$wday%%6 == 
    0, "weekend", "weekday")
# For Sunday and Saturday : weekend, Other days : weekday
activityFilled$day = factor(activityFilled$day, levels = c("weekday", "weekend"))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r}
stepsInterval2 = aggregate(steps ~ interval + day, activityFilled, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = stepsInterval2, aspect = 1/2, 
    type = "l")
```
